---
title: "Homework 1"
author: "Yuhao Mao 3170102264"
date: "2020/7/6"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*****
## **Problem 1: Iowa data set**

1. Load data with `read.csv()`.
```{r}
iowa.df<-read.csv("data/iowa.csv",header=T,sep=';')
```

2. Learn number of rows and columns.
```{r}
dim(iowa.df)
```
Therefore, there are 33 rows and 10 columns.

3. Get names of columns.
```{r}
colnames(iowa.df)
```

4. value in row 5, column 7.
```{r}
iowa.df[5,7]
```

5. Second row of the data set.
```{r}
iowa.df[2,]
```

*****

## Problem 2: **Syntax**

1. Explain following commands.

```{r, eval=FALSE}
vector1 <- c("5", "12", "7", "32") ## No print-out. 
## This sentence assigns the 
## combination/list to the variable vector1.

max(vector1) ## "7". 
## Because they are all strings and 
## the order of strings is based on characters at the beginning. 
## Here "7" is bigger than "5", "1" and "3".

sort(vector1) ## "12" "32" "5"  "7".
## Same as before, the strings are sorted
## based on characters at the beginning.

sum(vector1) ## Error.
## strings cannot be added.
```

2. Explain following commands.
```{r, eval=FALSE}
vector2 <- c("5",7,12)
vector2[2] + vector2[3] ## Error.
## string and integer cannot be added together.

dataframe3 <- data.frame(z1="5",z2=7,z3=12)
dataframe3[1,2] + dataframe3[1,3] ## 19.
## 7+12=19.

list4 <- list(z1="6", z2=42, z3="49", z4=126)
list4[[2]]+list4[[4]] ## 168.
## This indexing returns values. 42+126=168.
list4[2]+list4[4] ## Error.
## This indexing returns lists. Lists cannot be added together.
```

3. Producing lists and replicates.
  + Lists.
    From 1 to 10000 in increments of 372.
```{r}
seq(1, 10000, by=372)
```
    From 1 to 10000 with exactly 50 numbers.
```{r}
seq(1, 10000, length.out=50)
```

  + Replicates.
```{r}
rep(1:3, times=3)
```
```{r}
rep(1:3, each=3)
```
It is clear that the first one repeats the whole list for three times but the second one repeats each element for three times.

*****
## **Problem 3: Orings data set**
Read in data set.
```{r}
library(faraway)
data(orings)
```
Extract used rows
```{r}
Orings.used <- orings[c(1,2,4,11,13,18),]
Orings.used
```
Plot extracted information.
```{r, out.width=400, dig.align='center'}
plot(damage~temp, data=Orings.used, 
       xlab="temperature",ylab="damage",col="red", ylim=c(0,6),type='l')
```

Plot the whole data set.
```{r, out.width=400, dig.align='center'}
plot(damage~temp, data=orings, 
       xlab="temperature",ylab="damage",col="red", ylim=c(0,6),type='l')
```

*****
## **Problem 4: ais data set**
1. Read in data set.
```{r results="hide"}
library(DAAG,warn.conflicts=F, quietly=T)
data("ais")
```
Use `str()` to extract information.
```{r}
str(ais)
```
Whether it contains missing values.
```{r}
any(is.na(ais))
```
The answer is NO.

2. Create table to count the number of males and females for each sport.
```{r}
sport <- unique(ais$sport)
table.df <- read.table(text = "",col.names = sport)
for(i in 1:length(sport))
{
  s <-  sport[i]
  table.df[1,i] <-  sum(ais$sex=='f' & ais$sport==s)
  table.df[2,i] <-  sum(ais$sex=='m' & ais$sport==s)
}
row.names(table.df) = c('f', 'm')
table.df
```

Large imbalance sports:
```{r}
index <-  table.df[1,]/table.df[2,]>2 | table.df[1,]/table.df[2,]<0.5
colnames(table.df)[index]
```

*****
## Problem 5: Manitoba data set

Create data set:
```{r}
Manitoba.lakes <- setNames(
  data.frame(
    t(data.frame(c("Winnipeg", 217, 24387),
                 c("Winnipegosis", 254, 5374),
                 c("Manitoba", 248, 4624),
                 c("SouthernIndian", 254,2247),
                 c("Cedar", 253, 1353),
                 c("Island", 227, 1223),
                 c("Gods", 178, 1151),
                 c("Cross", 207, 755),
                 c("Playgreen", 217, 657)))
    ,row.names = NULL,stringsAsFactors = FALSE
  ), 
  c("district","elevation","area")
)
row.names(Manitoba.lakes) <- Manitoba.lakes$district
Manitoba.lakes$district <- NULL
Manitoba.lakes$elevation <-  as.numeric(Manitoba.lakes$elevation)
Manitoba.lakes$area <-  as.numeric(Manitoba.lakes$area)
Manitoba.lakes
```

1. Plot log2(area) against elevation.
```{r}
attach(Manitoba.lakes)
plot(log2(area) ~ elevation, pch=16, xlim=c(170,280))
# NB: Doubling the area increases log2(area) by 1.0
text(log2(area) ~ elevation, labels=row.names(Manitoba.lakes), pos=4)
text(log2(area) ~ elevation, labels=area, pos=2)
title("Relationship between log value of area and elevation\n among Manitoba’s Largest Lakes")
```

2. Plot area against elevation.
```{r}
plot(area ~ elevation, pch=16, xlim=c(170,280), ylog=T)
text(area ~ elevation, labels=row.names(Manitoba.lakes), pos=4, ylog=T)
text(area ~ elevation, labels=area, pos=2, ylog=T)
title("Manitoba’s Largest Lakes")
```

## **Problem 6: Dot chart for Manitoba's lakes.**

1. Areas on a linear scale.
```{r out.width=400}
dotchart(area,labels=row.names(Manitoba.lakes),xlab="area")
```


2. Areas on a logarithmic scale.
```{r out.width=400}
dotchart(log2(area),labels=row.names(Manitoba.lakes),xlab="log2(area)")
```

