---
title: "Homework 4"
author: "Yuhao Mao"
date: "2020/7/12"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r message=F, include=F}
library(tidyverse)
```

```{r message=F}
ckm_nodes <- read_csv('data/ckm_nodes.csv')
ckm_network <- read.table("data/ckm_network.dat")
```

## 1. Clean the data.
```{r}
ckm_nodes.valid.index <- which(!is.na(ckm_nodes$adoption_date))
ckm_nodes <- ckm_nodes[ckm_nodes.valid.index,]
ckm_network <- ckm_network[ckm_nodes.valid.index, ckm_nodes.valid.index]
```

## 2. Reformat data.
```{r}
data <- data.frame(month=rep(c(1:16,Inf),each=125),doctor=rep(as.numeric(rownames(ckm_nodes)),times=17),adoption_date=rep(ckm_nodes$adoption_date,times=17))
data$this_month <- data$month==data$adoption_date
data$before_this_month <- data$month>data$adoption_date

cal.before <- function(vec){
  doc <- vec[2]
  month <- vec[1]
  contact <- which(ckm_network[doc,]==1)
  contact_adopt <- ckm_nodes$adoption_date[contact]
  return(sum(contact_adopt<month))
}
cal.before_or_eq <- function(vec){
  doc <- vec[2]
  month <- vec[1]
  contact <- which(ckm_network[doc,]==1)
  contact_adopt <- ckm_nodes$adoption_date[contact]
  return(sum(contact_adopt<=month))
}

data$num_contact_before <- apply(data, 1, cal.before)
data$num_contact_before_or_eq <- apply(data, 1, cal.before_or_eq)
data$adoption_date <- NULL
```

Because there are 125 doctors and 17 month values, total number of rows is $125\times17=2125$.

## 3. Probabilities.
+a. 
```{r}
max(data[,c(5,6)])
```
Since the maximum value of k is 20, we can only estimate at most 21 values of k.

+b. Calculate $p_k$.
```{r}
pk <- c()
for(k in 0:20){
  joint <- sum(data$this_month&data$num_contact_before==k)
  cond <- sum(data$num_contact_before==k)
  pk <- c(pk, joint/cond)
}
plot(0:20, pk, type='l', xlab='k')
```

+c. Calculate $q_k$.
```{r}
qk <- c()
for(k in 0:20){
  joint <- sum(data$this_month&data$num_contact_before_or_eq==k)
  cond <- sum(data$num_contact_before_or_eq==k)
  qk <- c(qk, joint/cond)
}

plot(0:20, qk, type='l', xlab='k')
```

## 4. Interpretation of $p_k$.
+a. Linear model.
```{r}
pk <- data.frame(pk=pk,k=0:20)
pk <- pk %>% filter(!is.na(pk))
linear.result <- lm(pk~k, data=pk)
linear.result$coefficients
```
+b. Logistic model.
```{r}
logit.result <- glm(pk~k,data=pk,family = "binomial")
summary(logit.result)
```
Therefore, if $k$ increases by 1, then $a+bk$ decreases because its coefficient is negative and $e^{a+bk}$ is smaller. By the monotonicity of logit function, we find $p_k$ is smaller.

+c. Plot fitted values.
```{r}
pk %>% ggplot() + geom_line(aes(k, pk), col="black", size=1.5) +
  geom_line(aes(k, linear.result$fitted.values), col="blue", size=1.5) +
  geom_line(aes(k, logit.result$fitted.values), col="red", size=1.5)
```
I prefer the logistic model because it seems to capture the trend when k increases while the linear model exihibts some kinds of deviation.
